---
title: "Data Science and Machine Learning Bootcamp with R"
output: html_notebook
---

[Course dashboard](https://www.udemy.com/data-science-and-machine-learning-bootcamp-with-r/learn/v4/overview)

## Section 1 - Course Introduction

### Lecture 1 - Introduction to Course

Author - [Jose Portilla](http://www.pieriandata.com/)

Couse is suitable for:

- Developers who want to learn about Machine Learning
- Complete begginers who wan't to learn coding and ML

Course covers:

- R programming
- Machine Learning
- Data Visualization

Including:

- Basic R programming topics
- Data visualisation in R
- Linear regression
- Regression trees
- Random forests
- Neural networks
- Web Data Mining

Course includes:

- ML Portfolio projects
- R Programming Exercises
- Capstone Data Projects
- Over 100HD Video Lectures
- Code and Explanation Notebooks for each video
- Certificate of Completion

### Lecture 2 - Course Curriculum

Sections:

1. Course Introduction
    - what is Data Science
    - Course FAQs
2. Course Best Practices
    - How to Get Help in the Course
3. Windows Installation Set-Up
    - installation
4. Mac OS Installation Set-Up
    - installation
5. Linux Installation
    - installation
6. Development Environment Overview
7. Introduction to R Basics
    - basic R
8. R Matrices
    - basic R
9. R Data Frames
    - basic R
10. R Lists
    - basic R
11. Data Input and Output with R
    - basic R
    - CSV
    - Excel
    - SQL
    - Web Scraping
12. R Programming Basics
    - basic programming
        - funtions
        - loops
        - statements
    - data manipulation
13. Advanced R Programming
    - advanced programming
14. Data Manipulation with R
    - data handling and manipulation
15. Data Visualization with R
    - ggplot
16. Data Visualization Project
    - project
17. Interactive Visualizations with Plotly
    -Plotly
18. Capstone Data Project
    - project
19. Introduction to Machine Learning with R
20. Machine Learning with R - Linear Regression
21. Machine Learning Project - Linear Regression
22. Machine Learning with R - Logistic Regression
23. Machine Learning Project - Logistic Regression
24. Machine Learning with R - K Nearest Neighbors
25. Machine Learning Project - K Nearest Neighbors
26. Machine Learning with R - Decision Trees and Random Forests
27. Machine Learning Project - Decision Trees and Random Forests
28. Machine Learning with R - Support Vector Machines
29. Machine Learning Project - Support Vector Machines
30. Machine Learning with R - K-means Clustering
31. Machine Learning Project - K-means Clustering
32. Machine Learning with R - Natural Language Processing
33. Machine Learning with R - Neural Nets
34. Machine Learning Project - Neural Nets
35. Bonus Section - Discounts for Other Courses

### Lecture 3 - What is Data Science (skipped)

### Lecture 4 - Course FAQ (skipped)

1. Where can I download R and RStudio?
2. Where can I get the Course Notes?
3. Why are the Course Notes in HTML?
4. How do I get a certificate of Completion? 

## Section 2 - Course Best Practices

### Lecture 5 - How to Get Help in the Course! (skipped)

### Quiz 1: Welcome to the Course (skipped)

### Lecture 6 - Installation and Set-Up (skipped)

## Section 3 - Windows Installation Set-Up

### Lecture 7 - Windows Installation Procedure (skipped)

## Section 4 - Mac OS Installation Set-Up

### Lecture 8 - Mac OS Installation Procedure (skipped)

## Section 5 - Linux Installation

### Lecture 9 - Linux/Unbuntu Installation Procedure (skipped)

## Section 6 - Development Environment Overview

### Lecture 10 - Development Environment Overview

Overview of 2 following lessons

### Lecture 11 - Course Notes

How to get - resources section of the course

Most lessons have corresponding notebook containing code from the lecture

Notebooks contain:

- Exercises
- Answers to exercises
- Projects input data
- Projects output data
- Sample projects

### Lecture 12 - Guide to RStudio (skipped)

## Section 7 - Introduction to R Basics

[notebooks for this section](..\R-for-Data-Science-and-Machine-Learning\Basics of R and Data Types\R Basics)

### Lecture 13 - Introduction to R Basics

Structure of the notes

### Lecture 14 - Arithmetic with R

Modulo
```{r}
5 %% 2
```

### Lecture 15 - Variables

Variable naming in allows usage of

- downcase letters
- uppercase letters
- `_` underscore
- `.` dot

### Lecture 16 - R Data Types

To check class of object use `class()` function

### Lecture 17 - Vector Basics

How to construct vector using `c()` (combine) function and assign name to each element of the vector

#### Creating Vectors

Vector - One dimensional array that can hold character, numeric or boolean datatypes

If vector contains various datatypes on input, it truncates all data to one datatype

#### Assigning names to elements of the vector

```{r}
# creating vector
temps <- c(72,71,68,73,69,75,76)
temps
# asigning names
days <- c('Mon','Tue','Wed','Thu','Fri','Sat','Sun')
names(temps) <- days
temps
class(temps)
# accesing value of named element of the vector
temps['Wed']
```

### Lecture 18 - Vector Operations

`prod()` - gives product of all elements of the vector

[CRAN R Reference Card](https://cran.r-project.org/doc/contrib/Short-refcard.pdf)

### Lecture 19 - Vector Indexing and Slicing

Filtering
```{r}
v <- c(1,2,3,4)
names(v) <- c('a','b','c','d')
v
my.filter <- v>2
# accessing only elements > 2
v[v>2]
v[my.filter]
```

### Lecture 20 - Getting Help with R

Use `help()` function

E.g. `help('vector')`

`help.search()`  or `??<name>` looks for all related articles

### Lecture 21 - Comparing Operators

`> < >= <= == !=`

### Lecture 22 - R Basics Training Exercise

Explanation of usage of exercises in course notes

Exercises:

What is two to the power of five?
```{r}
2^5
```

Create a vector called stock.prices with the following data points: 23,27,23,21,34
```{r}
stock.prices <- c(23,27,23,21,34)
stock.prices
```

Assign names to the price data points relating to the day of the week, starting with Mon, Tue, Wed, etc...

```{r}
names(stock.prices) <- c('Mon','Tue','Wed','Thu','Fri')
stock.prices
```

What was the average (mean) stock price for the week? (You may need to reference a built-in function)

```{r}
mean(stock.prices)
```

Create a vector called over.23 consisting of logicals that correspond to the days where the stock price was more than $23

```{r}
over.23 <- stock.prices > 23
over.23
```

Use the over.23 vector to filter out the stock.prices vector and only return the day and prices where the price was over $23

```{r}
stock.prices[over.23]
```

Use a built-in function to find the day the price was the highest

```{r}
stock.prices[stock.prices == max(stock.prices)]
```

### Lecture 23 - R Basics Training Exercise - Solutions Walkthrough

## Section 8 - R Matrices

### Lecture 24 - Introduction to R Matrices

Matrices let us store data 2-dimensionally

### Lecture 25 - Creating a Matrix

Basics
```{r}
# creating a matrix
matrix(1:10,nrow=2)
# byrow parameter tells us if matrix should be filled up by row or by column
# F is default value
matrix(1:12, byrow = F, nrow = 3)
# TRUE
matrix(1:12, byrow = T, nrow = 3)
```

```{r}
# some made-up stock values
goog <- c(450,451,452,445,468)
msft <- c(230,231,232,233,220)
stocks <- c(goog,msft)
stocks 

# creating a matrix
stock.matrix <- matrix(stocks,byrow = T, nrow = 2)
stock.matrix

# naming columns and rows
# names() function is applicable, but we will use other functions


```
