---
title: "Data Science and Machine Learning Bootcamp with R"
output: html_notebook
---

[Course dashboard](https://www.udemy.com/data-science-and-machine-learning-bootcamp-with-r/learn/v4/overview)

## Section 1 - Course Introduction

### Lecture 1 - Introduction to Course

Author - [Jose Portilla](http://www.pieriandata.com/)

Couse is suitable for:

- Developers who want to learn about Machine Learning
- Complete begginers who wan't to learn coding and ML

Course covers:

- R programming
- Machine Learning
- Data Visualization

Including:

- Basic R programming topics
- Data visualisation in R
- Linear regression
- Regression trees
- Random forests
- Neural networks
- Web Data Mining

Course includes:

- ML Portfolio projects
- R Programming Exercises
- Capstone Data Projects
- Over 100HD Video Lectures
- Code and Explanation Notebooks for each video
- Certificate of Completion

### Lecture 2 - Course Curriculum

Sections:

1. Course Introduction
    - what is Data Science
    - Course FAQs
2. Course Best Practices
    - How to Get Help in the Course
3. Windows Installation Set-Up
    - installation
4. Mac OS Installation Set-Up
    - installation
5. Linux Installation
    - installation
6. Development Environment Overview
7. Introduction to R Basics
    - basic R
8. R Matrices
    - basic R
9. R Data Frames
    - basic R
10. R Lists
    - basic R
11. Data Input and Output with R
    - basic R
    - CSV
    - Excel
    - SQL
    - Web Scraping
12. R Programming Basics
    - basic programming
        - funtions
        - loops
        - statements
    - data manipulation
13. Advanced R Programming
    - advanced programming
14. Data Manipulation with R
    - data handling and manipulation
15. Data Visualization with R
    - ggplot
16. Data Visualization Project
    - project
17. Interactive Visualizations with Plotly
    -Plotly
18. Capstone Data Project
    - project
19. Introduction to Machine Learning with R
20. Machine Learning with R - Linear Regression
21. Machine Learning Project - Linear Regression
22. Machine Learning with R - Logistic Regression
23. Machine Learning Project - Logistic Regression
24. Machine Learning with R - K Nearest Neighbors
25. Machine Learning Project - K Nearest Neighbors
26. Machine Learning with R - Decision Trees and Random Forests
27. Machine Learning Project - Decision Trees and Random Forests
28. Machine Learning with R - Support Vector Machines
29. Machine Learning Project - Support Vector Machines
30. Machine Learning with R - K-means Clustering
31. Machine Learning Project - K-means Clustering
32. Machine Learning with R - Natural Language Processing
33. Machine Learning with R - Neural Nets
34. Machine Learning Project - Neural Nets
35. Bonus Section - Discounts for Other Courses

### Lecture 3 - What is Data Science (skipped)

### Lecture 4 - Course FAQ (skipped)

1. Where can I download R and RStudio?
2. Where can I get the Course Notes?
3. Why are the Course Notes in HTML?
4. How do I get a certificate of Completion? 

## Section 2 - Course Best Practices

### Lecture 5 - How to Get Help in the Course! (skipped)

### Quiz 1: Welcome to the Course (skipped)

### Lecture 6 - Installation and Set-Up (skipped)

## Section 3 - Windows Installation Set-Up

### Lecture 7 - Windows Installation Procedure (skipped)

## Section 4 - Mac OS Installation Set-Up

### Lecture 8 - Mac OS Installation Procedure (skipped)

## Section 5 - Linux Installation

### Lecture 9 - Linux/Unbuntu Installation Procedure (skipped)

## Section 6 - Development Environment Overview

### Lecture 10 - Development Environment Overview

Overview of 2 following lessons

### Lecture 11 - Course Notes

How to get - resources section of the course

Most lessons have corresponding notebook containing code from the lecture

Notebooks contain:

- Exercises
- Answers to exercises
- Projects input data
- Projects output data
- Sample projects

### Lecture 12 - Guide to RStudio (skipped)

## Section 7 - Introduction to R Basics

[notebooks for this section](..\R-for-Data-Science-and-Machine-Learning\Basics of R and Data Types\R Basics)

### Lecture 13 - Introduction to R Basics

Structure of the notes

### Lecture 14 - Arithmetic with R

Modulo
```{r}
5 %% 2
```

### Lecture 15 - Variables

Variable naming in allows usage of

- downcase letters
- uppercase letters
- `_` underscore
- `.` dot

### Lecture 16 - R Data Types

To check class of object use `class()` function

### Lecture 17 - Vector Basics

How to construct vector using `c()` (combine) function and assign name to each element of the vector

#### Creating Vectors

Vector - One dimensional array that can hold character, numeric or boolean datatypes

If vector contains various datatypes on input, it truncates all data to one datatype

#### Assigning names to elements of the vector

```{r}
# creating vector
temps <- c(72,71,68,73,69,75,76)
temps
# asigning names
days <- c('Mon','Tue','Wed','Thu','Fri','Sat','Sun')
names(temps) <- days
temps
class(temps)
# accesing value of named element of the vector
temps['Wed']
```

### Lecture 18 - 